\documentclass{article}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{float}
\usepackage{xcolor}
\usepackage{cite}
\usepackage{graphicx} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage[a4paper, margin=2cm]{geometry}

\title{SPRING Pseudo random functions}
\author{Jocelyn Fagard}
\date{March 2026}

\begin{document}

\maketitle

\section{Context}
Let $n,m \in \mathbb{N}^*$; let $q$ be a prime integer and consider the ring
\[
A := \mathbb{F}_q[X]/(X^m + 1)
\]
as well as $(n+1)$ polynomials $K:=(k_0, \dots, k_n) \in A^{n+1}$.
We also assume we have a rounding function $S : \mathbb{F}_q^m \to \{0,1\}^m$ which is the concatenation of a function $s : \mathbb{F}_q \to \{0,1\}$.

We are then interested in studying the function
\[
F_K : \{0,1\}^n \to \{0,1\}^m
\]
that maps $x := (x_1, \dots, x_n)$ to
\[
S\!\left( k_0 \prod_{i=1}^n k_i^{x_i} \right).
\] that we want to be pseudorandom.

\section{Shifted keys}
Let $k_0 \in A$, here we assume that for all $i \in \{0,\dots,n-1\}$ $k_{i+1}:= \sigma(k_i)$ where $\sigma$ is the cyclic shift  : $P \in \mathbb{F}_q[X]_{<m} \to X  P$ mod $(X^m-1)$.
Given a polynomial $r$ and an non-negative integer $l$ we note $r_l$ the $(l+1)$-th coefficient of $r$. Let's start with a remark.

\textbf{Claim.} Assume that $m$ is even.
Let $a,b \in A$. Then we have the following equality:
\[
(ab)_{\frac{m}{2} - 1}
=
- \left( \sigma^{\frac{m}{2}}(a)\, \sigma^{-\frac{m}{2}}(b) \right)_{\frac{m}{2} - 1}.
\]

\medskip

\textbf{Proof.}
This can be obtained by direct computation. We write
\[
(ab)_{\frac{m}{2} - 1}
=
\sum_{i=0}^{\frac{m}{2}-1} a_i \, b_{\frac{m}{2}-1 - i}
-
\sum_{i=\frac{m}{2}}^{m-1} a_i \, b_{\frac{m}{2}-1 - i},
\]
and
\[
\left( \sigma^{\frac{m}{2}}(a)\, \sigma^{-\frac{m}{2}}(b) \right)_{\frac{m}{2} - 1}
=
\sum_{i=0}^{\frac{m}{2}-1} a_{i-\frac{m}{2}} \, b_{\frac{m}{2}-1-i+\frac{m}{2}}
-
\sum_{i=\frac{m}{2}}^{m-1} a_{i-\frac{m}{2}} \, b_{\frac{m}{2}-1-i+\frac{m}{2}},
\]
where indices are taken modulo $m$.

By a change of variable, we obtain
\[
\left( \sigma^{\frac{m}{2}}(a)\, \sigma^{-\frac{m}{2}}(b) \right)_{\frac{m}{2} - 1}
=
\sum_{i=\frac{m}{2}}^{m-1} a_i \, b_{m-1-i}
-
\sum_{i=0}^{\frac{m}{2}-1} a_i \, b_{\frac{m}{2}-1-i}
=
- (ab)_{\frac{m}{2} - 1}.
\]
\hfill $\square$

\textbf{Consequence.}
Assume that $m$ is even and that $n \geq \frac{m}{2}$. Then, by the previous claim, we have
\[
(k_0 k_{\frac{m}{2}})_{\frac{m}{2}-1}
=
- \left( \sigma^{\frac{m}{2}}(k_0)\, \sigma^{-\frac{m}{2}}(k_{\frac{m}{2}}) \right)_{\frac{m}{2}-1}.
\]
Moreover, by the way the keys are generated, we also have
\[
 \left( \sigma^{\frac{m}{2}}(k_0)\, \sigma^{-\frac{m}{2}}(k_{\frac{m}{2}}) \right)_{\frac{m}{2}-1}
=
 (k_{\frac{m}{2}} k_0)_{\frac{m}{2}-1}.
\]
Hence,
\[
(k_0 k_{\frac{m}{2}})_{\frac{m}{2}-1} = 0.
\]

\medskip

Assume now that $s(0) = b \in \{0,1\}$. Then a simple algorithm to distinguish $F_K$ from a random function is the following:

\begin{itemize}
    \item Query the function on $x := (\delta_{i, \frac{m}{2}})_{1 \leq i \leq n}$.
    \item Let $y := (y_1, \dots, y_m)$ be the output.
    \item If $y_{\frac{m}{2}-1} = b$, output that the function is $F_K$; otherwise, output that it is random function.
\end{itemize}

The success probability of this distinguisher is
\[
\frac{1}{2} + \frac{1}{4} = \frac{3}{4}.
\]

\section{Linear relation between keys}
In this section, we assume that the keys $k_1, k_2, \dots, k_n$ are sampled independently and uniformly at random from a $d$-dimensional linear subspace of $\mathbb{F}_q[X]_{<m}$, with we identify with $\mathbb{F}_q^m$, where $1 \leq d \leq m$. We denote by $(e_1, \dots, e_d)$ a basis of this subspace.

We also fix an integer $1 \leq \ell \leq n$, which represents the Hamming weight $w_H(x)$ of the input $x := (x_1, x_2, \dots, x_n)$ used in our queries.

\medskip

Let
\[
E_\ell := \mathrm{span} \left\{ k_0 \prod_{i=1}^n k_i^{x_i} \;\middle|\; x \in \{0,1\}^n,\; w_H(x) = \ell \right\}.
\]
Then we have the inclusion
\[
E_\ell \subseteq \mathrm{span} \left\{ \prod_{i=1}^d e_i^{z_i} \;\middle|\; z_i \in \mathbb{N},\; \sum_{i=1}^d z_i = \ell \right\}.
\] Therefore, $E_\ell$ is generated by at most $\binom{d+\ell-1}{d -1}$ (which is the number of monomials of degree $\ell$ in d variables) elements. This provides an upper bound on its dimension, denoted $c_{d,\ell}$.

Now assume we take $x, y \in \{0,1\}^n$ uniformly at random among messages of Hamming weight $\ell$ with $x \neq y$.

We can then expect
\[
\Pr[F_K(x) = F_K(y)] = \frac{1}{q^{c_{d,\ell}}} \geq \frac{1}{q^{\binom{d+\ell-1}{d-1}}} = \frac{1}{2^{\log_2(q)\binom{d+\ell-1}{d-1}}} =: p_{d,\ell}.
\]
This probability is maximized for $\ell=1$, where it equals
\[
p_{d,1} = \frac{1}{2^{\log_2(q)\, d}}.
\]
Note that for a truly random function, this collision probability is $2^{-m}$.

Let's now present a Distinguisher.
\medskip

\textbf{Distinguisher:}
\begin{itemize}
    \item Take two different vectors of Hamming weight 1 and query the function on them.
    \item If there is a collision, output that the function is $F_K$; otherwise, output that it is either $F_K$ or a random function with probability $1/2$ each.
\end{itemize}

\medskip

The success probability of this distinguisher can then be expressed as
\[
\Pr[\text{success}] \geq \frac{1}{2} \cdot \frac{1}{2} \cdot \left(1 - 2^{-m}\right) 
+ \frac{1}{2} \left( p_{d,1} + (1 - p_{d,1}) \cdot \frac{1}{2} \right) = \frac{1}{2} + \frac{p_{d,1}}{4} - \frac{1}{2^{m+2}}.
\]
\textbf{Consequence.}  
The function $F_K$ is not pseudorandom if $p_{d,1}$ is the inverse of a polynomial in $m$, that is, if 
\[
d = \log_2 \bigl( U(m) \bigr)
\]
for some polynomial $U \in \mathbb{R}[X]$.

\textbf{Remark.}  
Fortunately, the probability that the keys lie in a low-dimensional space when they are sampled uniformly at random and independently is very low, since
\[
\Pr\bigl[\dim(\mathrm{span}(k_1, k_2, \dots, k_n)) = d \bigr]
= \frac{\prod_{i=0}^{d-1} \frac{(q^m - q^i)(q^n - q^i)}{q^d - q^i}}{q^{mn}}.
\]

\section{Freshman Dream}

In this section, we remark that no nontrivial ring $R$ exists such that for all $a,b,c,d \in R$, the equality
\[
(a+b)(c+d) = ac + bd
\]
holds.

To see this, suppose such a ring exists. Let $a \in R$, and denote by $0$ and $1$ the additive and multiplicative identities of $R$, respectively.

On the one hand, using the assumed property, we have
\[
(a+0)(1+0) = a \cdot 1 = a.
\]

On the other hand, using the assumed property and commutativity of addition, we get
\[
(a+0)(1+0) = (0+a)(1+0) = 0 \cdot 1 + a \cdot 0 = 0 + 0 = 0.
\]

Hence, $a = 0$ for all $a \in R$, which means that $R$ is the zero ring. Therefore, no nontrivial ring satisfies the given property.

\section{Some open problems about polynomial permutations}

Let $p$ be a prime number and let $d \geq 2$.

A polynomial permutation of $\mathbb{Z}_{p^d}$ is a polynomial 
$f \in \mathbb{Z}_{p^d}[X]$ that induces a permutation of the set $\mathbb{Z}_{p^d}$.

In previous works, it was proved that:

\vspace{0.3cm}

\textbf{Theorem.} 

A polynomial $f \in \mathbb{Z}_{p^d}[X]$ is a permutation polynomial 
if and only if its reduction modulo $p$ is a permutation polynomial of $\mathbb{Z}_p$ 
and $f'$ does not vanish on $\mathbb{Z}_p$.

\vspace{0.3cm}

In addition, the number of permutation polynomials of degree $n$ over $\mathbb{Z}_{p^d}$, 
denoted by $N(p^d,n)$, was left as an open question 
for $3 \leq n \leq 2p-2$.

Here, we address these questions.

\vspace{0.3cm}

\textbf{Lemma.} 

The number of permutation polynomials of $\mathbb{Z}_p$ 
of degree $p-1$ is zero.

Hence, for all $d \geq 2$, 
\[
N(p^d,p-1) = 0.
\]

\vspace{0.3cm}

\textbf{Proof.}

Given a permutation $\sigma$ of $\{0,1,\dots,p-1\}$, 
Lagrange interpolation shows that the unique polynomial 
of degree at most $p-1$ inducing this permutation is

\[
\sum_{i=0}^{p-1} \sigma(i) 
\prod_{\substack{0 \leq j \leq p-1 \\ j \neq i}} 
\frac{X - j}{i - j}.
\]

The coefficient of $X^{p-1}$ is therefore

\[
\sum_{i=0}^{p-1} \sigma(i) 
= \sum_{i=0}^{p-1} i 
= 0 \quad \text{in } \mathbb{Z}_p.
\]

Thus, no permutation polynomial over $\mathbb{Z}_p$ 
has degree exactly $p-1$.

\hfill $\square$

\vspace{0.3cm}
\textbf{Lemma.}

Let $p \geq 3$, we have:

\[
\frac{1}{2} \leq\frac{N(p, p-2)}{p!} \leq 1.
\]

That is at least half of polynomial permutations have degree $p-2$.

\textbf{Proof.}
Again, given a permutation $\sigma$ of $\mathbb{Z}_p$ using Lagrange interpolation formula, we have that the coefficient in front of the monom of degree $p-2$ is :

\[
- \sum_{i=0}^{p-1} \sigma(i) (\sum_{j \neq i} (-j)) = \sum_{i=0}^{p-1} \sigma(i)\cdot i = <id, \sigma>
\] 

where $<\cdot,\cdot>$ denotes the scalar product, $id = (0,1,\cdots,p-1)$ and $\sigma = (\sigma(0),\sigma(1),\cdots, \sigma(p-1))$. Hence $\sigma$ has degree $p-2$ if and only if it is not orthogonal to $id$.
\vspace{0.3cm}

Given $j \in \mathbb{Z}_p$ we define, $A_j :=\{\tau \text{ permutation} \mid <\tau, id> = j \}$ and $a_j:=\#A_j$. We first claim that $\forall i \neq j \in \mathbb{Z}_p^*$ , $a_i=a_j$ by remarking that $\forall \tau$ such that $<\tau,id> \neq 0$ there is exactly one permutation among the non zero co-linear permutations $\{ \lambda \cdot\tau \mid \lambda \in \mathbb{Z}_p^*\}$ in every $A_j$ for $j\neq0$.  
We denote this common value of these $a_i$'s with $i\neq 0$ by $b$.

\vspace{0.3cm}

Secondly, let's note $f_{i,j}: \tau \to (\tau_i,\tau_j) \tau$ the application that switches the i-th and the j-th values. Then we remark $<f_{i,j}(\tau), id> = <\tau,id> + (\tau(j) - \tau(i))(i-j) \neq <\tau,id>$ for $\tau$ a permutation. So, each element of some $A_l$ is associated to an element which is not in $A_l$. 
\vspace{0.3cm}

Now, we consider the graph where the vertices are the permutations and there is an edge between two permutations if one is obtained by applying $f_{0,1}$. In this graph every vertices are in couple because $f_{0,1} \circ f_{0,1} = id$. Each of the $a_0$ vertex in $A_0$ is associated to a distinct elements among the $(p-1)b$ which are not in $A_0$. So $(p-1)b \geq a_0$. And consequently the number of permutation polynomials of degree $p-2$ which is $(p-1)b$ represents at least half of permutations.

\vspace{0.3cm}

\hfill $\square$

\textbf{Proposition.}

Let $1 \leq n \leq p-1$ and $d \geq 2$. Then
\[
\frac{N(p^d,\leq p-1+n)}{p^{(d-1)(p+n)}}
= p! \left( 
p^n - \sum_{k=1}^{p} (-1)^{k-1} \binom{p}{k} \, p^{\max(n-k,0)}
\right).
\]

\vspace{0.3cm}

\textbf{Proof.}

Any polynomial $f \in \mathbb{Z}_{p}[X]$ of degree at most $p-1+n$ can be uniquely written in the form
\[
f(x) = (x^p - x) q(x) + \sigma(x),
\]
where $\sigma$ is a polynomial of degree at most $p-1$ and $q$ is a polynomial of degree at most $n-1$.

Moreover, $\sigma$ induces a permutation of $\mathbb{Z}_p$, and every such permutation arises uniquely in this way.

\vspace{0.2cm}

By the previous theorem, $f$ is a permutation polynomial over $\mathbb{Z}_{p^d}$ if and only if its derivative does not vanish on $\mathbb{Z}_p$. A direct computation shows that
\[
f'(x) \equiv -q(x) + \sigma'(x) ,
\]
so the condition becomes
\[
\forall i \in \mathbb{Z}_p, \quad \sigma'(i) \neq q(i).
\]

\vspace{0.2cm}

Fix a permutation polynomial $\sigma$. We want to count the number of polynomials $q \in \mathbb{Z}_p[X]$ of degree at most $n-1$ such that
\[
\forall i \in \mathbb{Z}_p, \quad q(i) \neq \sigma'(i).
\]

For each $i \in \mathbb{Z}_p$, define
\[
A_i := \left\{ q \in \mathbb{Z}_p[X]_{< n} \;\middle|\; q(i) = \sigma'(i) \right\}.
\]

We are therefore interested in the cardinality of the complement of $\bigcup_{i \in \mathbb{Z}_p} A_i$.

\vspace{0.2cm}

Using the inclusion--exclusion principle, and the fact that for any distinct 
$i_1,\dots,i_k \in \mathbb{Z}_p$, the intersection
\[
A_{i_1} \cap \cdots \cap A_{i_k}
\]
has cardinality $p^{n-k}$ if $1 \leq k \leq n$, and $1$ if $k \geq n$, we obtain
\[
\left|\bigcup_{i \in \mathbb{Z}_p} A_i\right|
= \sum_{k=1}^{p} (-1)^{k-1} \binom{p}{k} \, p^{\max(n-k,0)}.
\]

\vspace{0.2cm}

Hence, the number of admissible polynomials $q$ is
\[
p^n - \sum_{k=1}^{p} (-1)^{k-1} \binom{p}{k} \, p^{\max(n-k,0)}.
\]

\vspace{0.2cm}

Since there are $p!$ possible permutation polynomials $\sigma$ over $\mathbb{Z}_p$, we finally obtain
\[
N(p^d,\leq p-1+n)
= p! \left( 
p^n - \sum_{k=1}^{p} (-1)^{k-1} \binom{p}{k} \, p^{\max(n-k,0)}
\right).
\]

\hfill $\square$

\section{DDT of the Exponential Function over $\mathbb{F}_q^*$}

Let $q$ be an odd prime number and let $g$ be a generator of $\mathbb{F}_q^*$. 
We consider the function
\[
f : \mathbb{Z}_{q-1} \to \mathbb{F}_q^*, \quad x \mapsto g^x.
\]

For $\delta \in \mathbb{Z}_{q-1}$ and $\Delta \in \mathbb{F}_q$, we define
\[
\mathrm{DDT}(\delta, \Delta) := \# \left\{ x \in \mathbb{Z}_{q-1} \;\middle|\; f(x) + f(x+\delta) = \Delta \right\}.
\]

\medskip

\textbf{Proposition.}
\begin{itemize}
    \item If $\delta \neq \frac{q-1}{2}$ and $\Delta \neq 0$, then 
    \[
    \mathrm{DDT}(\delta, \Delta) = 1.
    \]
    \item If $\delta = \frac{q-1}{2}$, then 
    \[
    \mathrm{DDT}\!\left(\tfrac{q-1}{2}, 0\right) = q-1.
    \]
    \item In all other cases, $\mathrm{DDT}(\delta, \Delta) = 0$.
\end{itemize}

Particularly, $\mathrm{DDT}$ does not depend on the generator $g$.

\medskip

\textbf{Proof.}

First assume that $\delta = \frac{q-1}{2}$. Then
\[
(g^\delta)^2 = g^{q-1} = 1,
\]
so $g^\delta \in \{\pm 1\}$. Since $g$ is a generator of order $q-1$, we must have $g^\delta \neq 1$, hence $g^\delta = -1$.

Therefore, for all $x \in \mathbb{Z}_{q-1}$,
\[
f(x+\delta) + f(x) = g^{x+\delta} + g^x = g^x(1 + g^\delta) = 0.
\]
This shows that
\[
\mathrm{DDT}\!\left(\tfrac{q-1}{2}, 0\right) = q-1.
\]

\medskip

Now assume that $\delta \neq \frac{q-1}{2}$. Then $g^\delta \neq -1$, hence $1 + g^\delta \neq 0$. The map
\[
y \mapsto y(1 + g^\delta)
\]
is therefore a permutation of $\mathbb{F}_q^*$.

Thus, for any $\Delta \in \mathbb{F}_q^*$, the equation
\[
g^x + g^{x+\delta} = \Delta
\]
is equivalent to
\[
g^x (1 + g^\delta) = \Delta,
\]
which admits a unique solution $g^x \in \mathbb{F}_q^*$, and hence a unique $x \in \mathbb{Z}_{q-1}$.

This proves that $\mathrm{DDT}(\delta, \Delta) = 1$ for $\Delta \neq 0$, and $0$ otherwise.

\hfill $\square$

\textbf{Another formula for SPRING}.

Let us define
\[
R_q := \mathbb{F}_q[X]/(X^n + 1),
\]
where \(n\) is a power of two, and \(q\) is an odd prime such that \(q-1\) is divisible by \(2n\). In this setting, there exists a primitive \(2n\)-th root of unity \(\omega \in \mathbb{F}_q\). For example, one may take \(q=257\) and \(n=128\).

Using the isomorphism induced by the FFT and fixing a generator \(g\) of \(\mathbb{F}_q^\ast\), we represent the keys
\(k_0, k_1, \dots, k_m \in R_q^\ast\)
as elements of \(\mathbb{Z}_{q-1}^n\). We define
\[
K := (k_1 \mid k_2 \mid \cdots \mid k_m) \in \mathbb{Z}_{q-1}^{n \times m},
\]
where the columns correspond to the keys \(k_1,\dots,k_m\). With this notation, the first step of SPRING consists in computing the affine map
\[
c = (c_1,\dots,c_m) \in \{0,1\}^m \longmapsto Kc + k_0 \in \mathbb{Z}_{q-1}^n.
\]

We then apply coordinate-wise exponentiation to
\[
x := (x_1,\dots,x_n) = Kc + k_0 \in \mathbb{Z}_{q-1}^n,
\]
and obtain
\[
y := (g^{x_1},\dots,g^{x_n}) \in \mathbb{F}_q^n,
\quad \text{which we denote by } y = g^x.
\]

\vspace{0.3cm}

The inverse FFT step consists in applying the matrix
\[
\frac{1}{n}V, \quad \text{where } V := \bigl(\omega^{-2ij}\bigr)_{0 \leq i,j \leq n-1}.
\]
We obtain
\[
z := \frac{1}{n} V y.
\]

We then perform a pointwise multiplication with the vector
\[
\eta := (1, \omega^{-1}, \dots, \omega^{-(n-1)}),
\]
yielding
\[
u := z \odot \eta.
\]

Finally, we apply coordinate-wise rounding \(\lfloor \cdot \rceil\), defined by
\[
\lfloor a \rceil :=
\begin{cases}
0 & \text{if } 0 \leq a < \frac{q}{4} \text{ or } \frac{3q}{4} \leq a < q, \\
1 & \text{otherwise}.
\end{cases}
\]

One easily checks that \(\lfloor a \rceil = \lfloor -a \rceil\) for all \(a \in \mathbb{F}_q\).

Finally, given \(K\), \(k_0\), and \(c\), the SPRING function is written as
\[
F_{K,k_0}(c)
:=
\left\lfloor
\left(
\frac{1}{n} V \cdot g^{Kc + k_0}
\right)
\odot \eta
\right\rceil.
\]

\vspace{0.3cm}

\textbf{Definition}.

We say two sets of keys $(k_0,k_1,\cdots,k_m)$ and $(k'_0,k'_1,\cdots,k'_m)$ are equivalent and we denote it by $\sim$ if :
\[
\forall i, \exists \epsilon \in \{ 0, 1\} , k_i = k'_i + \epsilon \cdot (\frac{q-1}{2}, \frac{q-1}{2}, \cdots, \frac{q-1}{2}) 
\]

Each equivalent class is composed of $2^m$ sets of keys.

\vspace{0.3cm}

\textbf{Proposition}
Let $(K,k_0) \sim (K',k'_0)$, then $F_{K,k_0} = F_{K',k'_0}$.

\vspace{0.3cm}
\textbf{Proof.}
Let's consider $c\in \{0,1\}^m$.Because of the equivalence of sets of keys, we have :

\[
Kc + k_0 = K'c + k'_0 + \epsilon (\frac{q-1}{2}, \frac{q-1}{2}, \cdots, \frac{q-1}{2}) $ where $\epsilon \in \{ 0, 1\}.
\]

\vspace{0.3cm}

By remarking that : $g^{\epsilon (\frac{q-1}{2}, \frac{q-1}{2}, \cdots, \frac{q-1}{2})} = ((-1)^\epsilon,(-1)^\epsilon,\cdots,(-1)^\epsilon) $ we obtain:

\[
F_{K,k_0}(c)
=
\left\lfloor
\left(
\frac{1}{n} V \cdot g^{Kc + k_0}
\right)
\odot \eta
\right\rceil = \left\lfloor
(-1)^\epsilon\left(
\frac{1}{n} V \cdot g^{K'c + k'_0}
\right)
\odot \eta
\right\rceil = F_{K',k'_0}(c).
\]

\hfill $\square$

\vspace{0.3cm}

\section{Some thoughts about linear forms in $\mathbb{Z}_{q-1}^n$}

In this section we will have $q-1:= 2^{\alpha}$.

\vspace{0.3cm}

We have seen sampling the key was equivalent to choose uniformly at random and independently a matrix $K \in \mathbb{Z}_{q-1}^{n \times m}$ as well as an offset vector $k_0 \in \mathbb{Z}_{q-1}^n$.

\vspace{0.3cm}

We are interested in this section to get the bias of the first function $g$, of SPRING:
\[
 c:=(c_1,\cdots,c_m) \in \{0,1\}^m \to g^{K \cdot c + k_0} \in (\mathbb{F}_q^*)^n.
\]

\vspace{0.3cm}

\textbf{Lemma.}
Let $v:=(v_1,v_2,\cdots,v_n)\in \mathbb{Z}_{q-1}^n$.
A product scalar linear form : $<v, \cdot >: \mathbb{Z}_{q-1}^n \to \mathbb{Z}_{q-1}$ is equidistributed if and only if there exist $i$ such that $v_i \in \mathbb{Z}_{q-1}^*$.

\vspace{0.3cm}

\textbf{Proof.}
If it was not the case then every coordinate of $v$ would be divisible by 2 and the image of $<v,\cdot>$ would be included in the set of even numbers, so not equidistributed.

\vspace{0.3cm}

Now without loss of generality, if $v_n$ is invertible, the map $l \in \mathbb{Z}_{q-1} \to l\cdot v_n \in \mathbb{Z}_{q-1}$ is a bijection. So  for every value $j \in \mathbb{Z}_{q-1}$ and every truncated vector $(x_1,\cdots,x_{n-1}) \in  \mathbb{Z}_{q-1}^{n-1} $ there is exactly one value $l \in \mathbb{Z}_{q-1}$ such that $<v, (x,l)> = j$. 

\hfill $\square$

\vspace{0.3cm}

\textbf{Definition.}
A family $a:=(a_1,\cdots,a_{\alpha}) \in \mathbb{Z}_{q-1}^{\alpha}$ is a binary generator of $\mathbb{Z}_{q-1}$ if:
\[
\{\sum_{i=1}^{\alpha} \epsilon_i \cdot a_i \mid \epsilon_i \in \{0,1\}\} = \mathbb{Z}_{q-1}.
\]

\vspace{0.3cm}

\textbf{Lemma.}
$a:=(a_1,\cdots,a_{\alpha})$ is a binary generator if and only if each $a_i$ has different 2-valuation  from the others.

\vspace{0.3cm}

\textbf{Proof.} 
For the sufficient condition we can assume without loss of generality that $v_2(a_i)=i-1$, that is $a_i = 2^i u_i$ whit $u_i$ odd.
Let $(\epsilon_i)_i, (\epsilon'_i)_i \in \{0,1\}^{\alpha}$ such that:
\[
\sum_{i=1}^{\alpha} \epsilon_i \cdot a_i =\sum_{i=1}^{\alpha} \epsilon'_i \cdot a_i
\]
that is:
\[
\sum_{i=1}^{\alpha} (\epsilon_i -\epsilon'_i)\cdot 2^i u_i =0 
\].

Now because this equation is modulo $q-1=2^{\alpha}$, it must be divisible by 2. So $\epsilon_1 -\epsilon'_1=0$ that is $\epsilon_1 =\epsilon'_1$. Because it is divisible by $2^2$ we must have $\epsilon_2 =\epsilon'_2$. And so on until the fact that it is divisible by $2^{\alpha -1}$ which implies $\epsilon_{\alpha} =\epsilon'_{\alpha}$.

\vspace{0.3cm}

For the necessary condition, assume the $(a_i)_i$ are ordered by increasing valuations, let $\eta$ be the least element such that $v_2(a_{\eta}) \neq \eta -1$. If $ v_2(a_{\eta}) \geq \eta$, then $(2 ^{\eta-1} - 1)\cdot 2^{\alpha +1 -\eta}$ binary linear combination can lead to integers with valuation $\leq \eta-1$, but there are $(2^{\eta}-1) 2^{\alpha-\eta}$ such numbers.  If $ v_2(a_{\eta}) \leq \eta -2 $ then $a_{\eta-1}, a_{\eta}$ have valuation $\eta-2$ and we can obtain twice more numbers of valuation $\eta -2$ with binary linear combinations as usual.

\hfill $\square$

\vspace{0.3cm}

\textbf{Corollary.}
Let's consider $v \in \mathbb{Z}_{q-1}^n$, that contains a binary generator of $\mathbb{Z}_{q-1}$ in its coordinates, then the function:
\[
c:=(c_1,\cdots,c_m) \in \{0,1\}^m \to <v,c>
\]
is equidistributed over $\mathbb{Z}_{q-1}$.

\vspace{0.3cm}

\textbf{Proof.} The proof is similar than the one for the first lemma. \hfill $\square$

\vspace{0.3cm}

\textbf{Theorem.}
Let's consider an integer $n \geq 1$ and $q$ a prime number such that : $q-1 := 2^{\alpha}$.
We  consider $v_1,\cdots, v_n \in (\mathbb{Z}_{q-1})^{\alpha n}$ such that each $v_j :=(v_{j,1},\cdots, v_{j,\alpha n)}$ has a support $I_j$ of $\alpha$ indices for which $ \{\textbf{v}_2(v_{j,i}) \mid i \in I_j \} = \{1,2,3,\cdots,\alpha-1\}$. In addition to this we assume:
\[
\forall i \neq j, I_i \cap I_j = \emptyset.
\]
Let's denote by $V \in \mathbb{Z}_{q-1}^{n \times n\alpha}$ the matrix with rows $(v_j)_j$. We have: 

If $b:=(b_1,\cdots,b_{\alpha n}) \sim U(\{0,1\}^{n \alpha})$ then $V \cdot b \sim U(\mathbb{Z}_{q-1}^n)$.

\vspace{0.3cm}

\textbf{Proof.}
Let's note $Y_i := \langle v_i, b\rangle $ the i-th coordinate of $V \cdot b$. Thanks to the formula of conditional probabilities, it would be sufficient to show that $Y_{l+1}$ given $Y_1,\cdots, Y_l$ is uniformly distributed over $\mathbb{Z}_{q-1}$ for all $l\in \{1,2,\cdots , n-1\}$.

\vspace{0.3cm}

Let's make a recursion. Assume $Y_1=y_1$ is fixed, then because of the previous Corollary, the set $(b_{i})_{i \notin I_1}$ is uniformly distributed because $(v_{1,i})_{i \in I_1}$ is a binary generator of $\mathbb{Z}_{q-1}$. And because of the fact $I_2 \cap  I_1 = \emptyset$, this implies $(b_{1,i})_{i \in I_2}$ are uniformly distributed and so $Y_2$ is also uniformly distributed knowing  $(v_{2,i})_{i \in I_2}$ is a binary generator.

\vspace{0.3cm}

For the recursion step, assume given $Y_1 = y_1,\cdots, Y_{l-1} = y_{l-1}$ fixed, the $(b_i)_{i \notin (I_1\cup I_2 \cdots \cup I_{l-1})}$ are uniformly distributed and so $Y_l$ is uniformly distributed. Then let's fix $Y_l = y_l$ then because $I_l \cap (I_1\cup I_2 \cdots \cup I_{l-1}) = \emptyset$, $(b_i)_{i \notin (I_1\cup I_2 \cdots \cup I_{l-1} \cup I_l)}$ is uniformly distributed and so in particular $I_{l+1} \cap (I_1\cup I_2 \cdots \cup I_{l}) = \emptyset$ leads to the fact that $Y_{l+1} \mid Y_1,\cdots, Y_l$ is uniform over $\mathbb{Z}_{q-1}$.

\hfill $\square$

\vspace{0.3cm}

\textbf{Corollary SPRING.}
In SPRING, if $m=\alpha n$ and if the key matrix  $K \in \mathbb{Z}_{q-1}^{n \times n \alpha}$ is chosen such that its rows $v_1,\cdots,v_n$ respect the 'two-valuation assumption' of the previous theorem. Then, noting:

\[
 G: b:=(b_1,\cdots,b_m) \in \{0,1\}^m \to g^{K \cdot b + k_0} \in (\mathbb{F}_q^*)^n.
\]

we have $G \cdot b \sim U((\mathbb{F}_q^*)^n)$.

\vspace{0.3cm}

\textbf{Proof.} \hfill $\square$

\vspace{0.3cm}
Now we are interested in looking at the second part of the SPRING where our vector $y:=g^{Kc+k_0}$ that we hope to be uniformly distributed over $(\mathbb{F}_q^*)^n$ will be multiplied by the matrix $V \in (\mathbb{F}_q^*)^{n \times n}$.

\vspace{0.3cm}

\textbf{Proposition.}
Let $n \geq 1$, let $v \in (\mathbb{F}_q^*)^n$, and consider the linear form $<v,\cdot>$ on $(\mathbb{F}_q)^n$ restricted to $(\mathbb{F}_q^*)^n$. For any $i \in \mathbb{F}_q$ we denote:
\[
a_i:= \#\{x \in (\mathbb{F}_q^*)^n \mid <v,x>=i\}.
\]

Then for all $i,j \neq 0$, $a_i = a_j=:b$ and we have $a_0 - b = (-1)^n$. That is the restriction of the linear form $v$ to $(\mathbb{F}_q^*)^n$ is nearly equidistributed.

\vspace{0.3cm}

\textbf{Proof.}
Firstly in each line $\mathbb{F}_q^* \cdot x$ with $x \in (\mathbb{F}_q^*)^n$ not orthogonal to $v$ ($<v,x> \neq 0$), there is exactly one vector that contributes to the $a_i$ $i\neq0$. Hence for all $i, j \neq 0$, $a_i = a_j =:b$.

\vspace{0.3cm}

So as to make some induction we note $x_n:=a_0$ and $y_n:=b$. Now we take $v_{n+1} \in \mathbb{F}_q^*$ so as to complete $v$, and by writing: $\forall y:=(x,x_{n+1})=(x_1,...,x_n,x_{n+1}) \in (\mathbb{F}_q^*)^{n+1}$, $<(v,v_{n+1}),y> = <v,x> + x_{n+1}v_{n+1}$ and noticing the multiplication by $v_{n+1}$ realizes a permutation of the set $\mathbb{F}_q^*$ we deduces the following induction formulas:

\[
x_{n+1} = (q-1) y_n
\]
\[
y_{n+1} = x_n + (q-2) y_n.
\]

that is: 

\[
\begin{pmatrix}x_{n+1} \\ y_{n+1}\end{pmatrix} = M \begin{pmatrix}x_{n} \\ y_{n}\end{pmatrix}
\]


with

\[
M:=\begin{pmatrix}0 & q-1 \\1 & q-2\end{pmatrix}.
\]

Now we can diagonalize it, this leads to:

\[
\frac{1}{q}PMP=\begin{pmatrix}q-1 & 0 \\0 & -1\end{pmatrix}
\]

with:
\[
P:=\begin{pmatrix} 1&q-1\\1&-1 \end{pmatrix}.
\]

Now so as to conclude, we remark $x_1 = 0, y_1=1$.

\[
\begin{pmatrix}x_{n} \\ y_{n}\end{pmatrix} = \frac{1}{q}P\begin{pmatrix}(q-1)^{n-1} & 0 \\0 & (-1)^{n-1}\end{pmatrix} P \begin{pmatrix}0 \\ 1\end{pmatrix}
\]
\[
\begin{pmatrix}x_{n} \\ y_{n}\end{pmatrix} = \frac{1}{q}\begin{pmatrix}(q-1)^n + (-1)^n(q-1) \\ (q-1)^n + (-1)^{n+1}\end{pmatrix}.
\]

We get: $a_0-b = x_n - y_n =(-1)^n$

\hfill $\square$

\textbf{Lemma.} Let's consider $\phi$ be a primtive $n$-th root of unity.
Let $j \in \{1,\cdots, n-1\}$, and let $\phi_j := (\phi^{ji})_{i \in \{0,\cdots,n-1\}}$, $id := (1,1,\cdots,1)$. Then:
\[
\langle id, \phi_j \rangle = 0.
\]

\vspace{0.3cm}

\textbf{Proof.}
This comes from the fact that for $j \in \{1,\cdots,n-1\}$, $\phi^j \neq 0$ and so:
\[
\sum_{i=0}^{n-1} \phi^{ji} = \frac{\phi^{jn} -1}{\phi^j - 1} = 0.
\]

\hfill $\square$

\vspace{0.3 cm}

\textbf{Hoping Lemma.}
Let $n < q$, and let $H$ be a subspace of $\mathbb{F}_q^n$ that contains $id$. In addition to this assume that $v \in \mathbb{F}_q^n$ is, not orthogonal to the whole space $H$ as well as $\langle v, id\rangle = 0$. Then:

\[
\langle v, \cdot \rangle : (\mathbb{F}_q^*)^n \cap H \to \mathbb{F}_q
\] 

is almost equidistributed over $\mathbb{F}_q$.

\vspace{0.3cm}

\textbf{Proof.}
First of all, we note $K_i:=\{x\in (\mathbb{F}_q^*)^n \cap H \mid \langle v, x\rangle\ = i\}$ for all $i \in \mathbb{F}_q$ and we remark by multiplying by a scalar that $\#K_i = \#K_j$ for $i,j \neq 0$. We denote this cardinal by $k_{\neq0}$. We also define $k_0:=\#K_0$. The goal is to show: $\mid k_0 - k_{\neq 0}\mid \leq 1$. We have already seen this result was true for $H:=\mathbb{F}_q^n$ that is $\text{dim}H = n$, as well as $v\in (\mathbb{F}_q^*)^n$. If $\text{dim}H = 1$, that is $H:= \mathbb{F}_q \cdot u$ for some vector $u$, then either $u \in (\mathbb{F}_q^*)^n$ in this case $k_0 = 0$ and $k_{\neq 0} = 1$. Or, if it is not the case, then $k_{\neq 0} = k_0= 0$.

\vspace{0.3cm}

For $i \in \mathbb{F}_q$, let $A_i:= \{x\in (\mathbb{F}_q^*)^n \cap H \mid \langle v, x\rangle\ = i\}$ and $a_i = \#A_i$. Because $id \in H$, for all $x \in H$,  $ \mathbb{F}_q \cdot id + x$ is included in $H$. And because $\langle v, id\rangle = 0$, $\langle v, \cdot\rangle$ is constant on $ \mathbb{F}_q \cdot id + x$ equal to $\langle v, x\rangle$.

\vspace{0.3cm}

Also for $i \in \mathbb{F}_q$, we define $M_i := \{x\in H \mid i \notin x \}$ the vectors in $H$ that does not have $i$ in its coordinates. For instance $M_0 = (\mathbb{F}_q^*)^n \cap H$. More generally given $I \subset \{1,\cdots,n\}$, we define $M_I:= \bigcap_{i \in I} M_i$ . Because $q < n$, $H= M_0 \cup M_1 \cup \cdots \cup M_{q-1}$. 

\vspace{0.3 cm}
Let $[x]:=x + \mathbb{F}_q  \cdot id \in H / (\mathbb{F}_q \cdot id)$  with $\langle v,x\rangle =: k$, let $I([x]) \subset \{1,\cdots,n\}$ be a biggest subset such that $x \in M_I$. Then the line $x + \mathbb{F}_q \cdot id$ will contribute $J([x]):=\#I([x])$ times to $a_k$. Because a non zero linear form is equidistributed over $\mathbb{F}_q$ we must have for all $k,l \in \mathbb{F}_q$:

\[
\sum_{[x], \langle v,x \rangle = k} J([x]) = \sum_{[x], \langle v,x \rangle = l} J([x])
\]

\vspace{0.3cm}

\hfill $\square $
\end{document}
